---
title: "Use Reactive Expression to Build Cacheable Function"
author: "Marlin JL.M"
date: "February 7, 2017"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Reactive Expression

If you have ever learned or used shiny, you will know about reactive expression.
In shiny app, we use reactive expressions to cache intermediate results and know
which parts to update when input has changed, so that the app can run more efficiently.

But in fact, reactive expression can not only be used by shiny app. Here I would
demonstrate how to build a function which caches its intermediate results
using reactive expression.


## Properties of a Reactive Expression

```{r, purl=FALSE}
library(shiny)
```

Here

```{r, purl=FALSE}
set.seed(42)
re <- reactiveValues()
re$x <- 42
re.a <- reactive( re$x + runif(1) )
re.b <- reactive( re.a() + 1 )

# Calculate the reactive expression (also cache the results in each step)
isolate(re.b())
```

If the reactive value that the expressions depend on, namely `re$x`, do not change,
next call to `isolate` will use the cached value.
Hence, the following result is the same as the previous,
despite there is a `runif` command.

```{r, purl=FALSE}
isolate(re.b())
```

Change the reactive value will recalculate the reactive expressions.

```{r, purl=FALSE}
re$x <- 69
isolate(re.b())
```


# A Cacheable Function


In the case that we want a function that cache its result, we can use a closure
that stores its reactive expressions in its parent environment.

```{r, purl=FALSE}
gen.f <- function () {
    # The place to store x (as a reactive source)
    reactv <- reactiveValues()
    # Store the reactive expressions
    re.x <- reactive({
        print("Getting re.x()...")
        # Calculating this is a little hard, take a rest first.
        Sys.sleep(1)
        reactv$x + 1
    })
    re.y <- reactive({
        print("Getting re.y()...")
        Sys.sleep(1)
        re.x() + 1
    })
    # Return a function (closure) that assign value of its argument as "reactiveValues"
    # in the parent environment and get the result of the reactive expressions.
    function (value) {
        reactv$x <<- value
        isolate(re.y())
    }
}
f <- gen.f()
```

Then let's examine this returned function.

1. The definition

```{r, purl=FALSE}
f
```

2. First run

```{r, purl=FALSE}
system.time(ans <- f(42))
ans
```

3. Second run with same argument

Run with the same argument will save some computing time by using
the cached result.

```{r, purl=FALSE}
system.time(ans <- f(42))
ans
```

4. Third run with a different argument

```{r, purl=FALSE}
system.time(ans <- f(69))
ans
```

# Generic Builer


```{r}
dots <- function(...) {
  eval(substitute(alist(...)))
}
isTheListNamed <- function (l) {
    nms <- names(l)
    if (any(is.null(nms)) || any(nms == ""))
        return(FALSE)
    else
        return(TRUE)
}
```


```{r}
#' Build Cacheable Functions
#' 
#' \code{reactFunc} is a function generator that returns a function that is cacheable when
#' combined with the use of \code{reactArg}.  See example.
#'
#' @name reactFunc
#'
#' @param ARGV
#'     A named list or vector that represent the formal arguments of the returned function.
#' @param ... 
#'     Name-expression pairs that describe the reactive expressions defined in the parent
#'     environment of the returned function. The last one will be used as the returned value
#'     of the returned function.  See example.
#' @return
#'     \code{reactFunc} returns a closure (function); \code{name} should be used inside the
#'     \code{...} arguments to describe the argument as a reactive source.
#' @export
#' @import shiny
#' @importFrom pryr make_function
#' @examples
#' ## Build
#' rf <- reactFunc(
#'     ARGV = alist(x = 42, y = ),
#'     x = reactArg(x),
#'     y = reactArg(y),
#'     a = {
#'         print("Getting a()...")
#'         Sys.sleep(0.5)
#'         x() + 1
#'     },
#'     b = {
#'         print("Getting b()...")
#'         Sys.sleep(0.5)
#'         y() + 1
#'     },
#'     ans = {
#'         print("Getting ans()")
#'         Sys.sleep(0.5)
#'         a() + b()
#'     }
#' )
#' ## Properties
#' #1. Definition
#' rf
#' #2. First run
#' m <- 6; n <- 9
#' system.time(ans <- rf(x = m, y = n))
#' ans
#' #3. Seconde run with the same arguments
#' system.time(ans <- rf(x = m, y = n))
#' ans
#' #4. Third run with an updated argument
#' n <- 7
#' system.time(ans <- rf(x = m, y = n))
#' ans
#' #5. Change the value of `x` to default
#' system.time(ans <- rf(y = n))
#' ans
reactFunc <- function (ARGV, ...) {
    reactContexts <- dots(...)
    reactNames <- names(reactContexts)
    arglist <- as.pairlist(ARGV)
    
    if (!isTheListNamed(reactContexts))
        stop("Names of the ... arguments must be specified.")
    if (!isTheListNamed(arglist))
        stop("ARGV must be a named list or vector.")
    
    ## The reactive expressions are defined here
    for (i in seq_along(reactContexts)) {
        assign(reactNames[[i]], shiny::reactive(reactContexts[[i]], quoted = TRUE))
    }
    
    ## The returned function will assign its arguments to this
    ## handler to make their values reactive.
    ReactiveSources <- shiny::reactiveValues()
    
    pryr::make_function(
        args = arglist,
        body = quote({
            assignedArgs <- #c(as.list(environment()), list(...))
                            as.list(environment())
            assignedArgNames <- names(assignedArgs)

            for (i in seq_along(assignedArgs)) {
                do.call(
                    what = `[[<-`,
                    args = list(
                        eval(parse(text = "ReactiveSources")),
                        assignedArgNames[[i]],
                        eval(parse(text = assignedArgNames[[i]]))
                    ),
                    envir = parent.env(environment())
                )
            }
            
            # Get result that produced by reactive expressions
            lastExprName <- tail(reactNames, 1)
            result <- shiny::isolate(
                do.call(lastExprName, args = list())
            )
            result
        })
    )
}
```

```{r}
## We need this function because `reactiveValues` can not be copied by hand:
## > r <- reactiveValues(x = 34)
## > x <- r$x     # Not copyable
#' @rdname reactFunc
#' @param name
#'     A symbol that corresponds to one of the names of \code{ARGV} (one of the argument name).
#'     See example.
#' @export
reactArg <- function (name) {
    namechr <- deparse(substitute(name))
    intended.expr <- parse(
        text = paste0("ReactiveSources$",namechr)
    )
    eval(intended.expr, envir = parent.frame())
}
```


# Example

The following is a simple example of creation of cacheable function using `reactFunc`.

```{r, purl=FALSE}
rf <- reactFunc(
    # `x` will have default value, `y` not.
    ARGV = alist(x = 42, y = ),
    x = reactArg(x),
    y = reactArg(y),
    a = {
        print("Getting a()...")
        Sys.sleep(0.5)
        x() + 1
    },
    b = {
        print("Getting b()...")
        Sys.sleep(0.5)
        y() + 1
    },

    ans = {
        print("Getting ans()")
        Sys.sleep(0.5)
        a() + b()
    }
)
```

Again let's look at this function `rf`.

1. Definition

```{r, purl=FALSE}
rf
```

2. First run

```{r, purl=FALSE}
m <- 6; n <- 9
system.time(ans <- rf(x = m, y = n))
ans
```

3. Seconde run with the same arguments

```{r, purl=FALSE}
system.time(ans <- rf(x = m, y = n))
ans
```

4. Third run with an updated argument

```{r, purl=FALSE}
n <- 7
system.time(ans <- rf(x = m, y = n))
ans
```

5. Change the value of `x` to default

```{r, purl=FALSE}
system.time(ans <- rf(y = n))
ans
```



# Use case

I think this kind of function can be most useful with shiny apps.
Of course you can choose to wrap all these reactive logic inside the shiny server,
but you can also separate each part in function which have a simple api and cached
intermediate results.

## TODO:
1. Include a use case of cacheable function.
2. How to copy/clone a function?




# Session

```{r, purl=FALSE}
sessionInfo()
```















```{r, eval=FALSE, include=FALSE, purl=FALSE}
######  ARCHIVE     #####-----------------------------------------


r <- reactiveValues(x = 234)
x <- r$x
delayedAssign("x", r$x)
ob <- reactive(x)
isolate(ob())
r$x <- 42
delayedAssign("x", r$x)
isolate(ob())


parseName <- function(name) {
    chr.name <- deparse(substitute(name))
    intended.expr <- parse(
        text = paste0("r$",chr.name)
    )
    eval(intended.expr, envir = parent.frame())
}
r <- reactiveValues(x = 2344)
wtf <- r$x
r1 <- reactive(r$x)
r2 <- reactive(parseName(x))
r1
r2
isolate(r1())
isolate(r2())
r$x <- 42
isolate(r1())
isolate(r2())

r.con <- alist(ind.x = x + 69, ind.y = )
r.nam <- names(r.con)



assignsomethinginparent <- function (v) {
    for (el in v)
        assignInParent(el, el)
}
assignsomethinginparent(c('wtf','eww','io'))


### Assign Reactive Expressions in a Loop
###  See: https://github.com/rstudio/shiny/issues/532
#1 This does not work
out1 <- vector("list", 3)
for(i in 1:3) {
  out1[[i]] <- reactive(i)
}
isolate(out1[[1]]())
#2 Either not work
out2 <- vector("list", 3)
for(i in 1:3) {
    local({
        out2[[i]] <<- reactive(i)
    })
}
isolate(out2[[1]]())
#3 Works
out3 <- vector("list", 3)
for(i in 1:3) {
    local({
        e <- i
        out3[[i]] <<- reactive(e)
    })
}
isolate(out3[[1]]())
#4 Works
out4 <- lapply(1:3, function(i) {
    reactive(i)
})
isolate(out4[[1]]())



context <- alist(x = 42,
                 y = x() + 1,
                 z = y() - 42)
name <- names(context)
for (i in 1:3) {
    assign(name[[i]], reactive(context[[i]] ,quoted = TRUE))
}
isolate(x())
isolate(y())
isolate(z())




f.gen <- function() {
    x <- 43
    function (v) {
        x <<- x + v
        x
    }
}
f <- f.gen()
f
f(1)
```


